---
title: "In the native speaker's eye_code"
output: 
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
  html_notebook: default
---

## Packages
```{r}
#install.packages('tidyverse')
#install.packages('viridis')
#install.packages('viridisLite')
#install.packages('markdown')
#install.packages('ggplot2')
#install.packages('dplyr')
#install.packages('lme4')
#install.packages('lmerTest')
#install.packages('tidyr')
#install.packages('effects')
#install.packages('Hmisc')
#install.packages("broom.mixed")

library(tidyverse)
library(tidyr)
library(viridis)
library(viridisLite)
library(markdown)
library(ggplot2)
library(dplyr)
library(lme4)
library(lmerTest)
library(effects)
library(broom.mixed)
```
## Data import
```{r}
#Eye = read.csv2("Eye.csv")
```
## Subsets
```{r}
df_s <- subset(Eye, IA_CAT == "S")
df_v <- subset(Eye, IA_CAT == "V")
df_a1 <- subset(Eye, IA_CAT == "A1")
df_o <- subset(Eye, IA_CAT == "O")
df_a2 <- subset(Eye, IA_CAT == "A2")
# List of subsets during first-pass reading
df_s_first <- subset(Eye, IA_CAT == "S" & IA_FIRST_FIX_PROGRESSIVE == "1")
df_v_first <- subset(Eye, IA_CAT == "V" & IA_FIRST_FIX_PROGRESSIVE == "1")
df_a1_first <- subset(Eye, IA_CAT == "A1" & IA_FIRST_FIX_PROGRESSIVE == "1")
df_o_first <- subset(Eye, IA_CAT == "O" & IA_FIRST_FIX_PROGRESSIVE == "1")
df_a2_first <- subset(Eye, IA_CAT == "A2" & IA_FIRST_FIX_PROGRESSIVE == "1")
```
## Contrast coding
```{r}
#Subjekt1
df_s_first$LENGTH <- as.factor(df_s_first$LENGTH)
df_s_first$GRAM <- as.factor(df_s_first$GRAM)
class(df_s_first$LENGTH)
class(df_s_first$GRAM)
levels(df_s_first$LENGTH)
levels(df_s_first$GRAM)
contrasts(df_s_first$LENGTH) <- c(0.5,-0.5)
contrasts(df_s_first$GRAM) <- c(-0.5,0.5)
```

```{r}
#Subjekt2
df_s$LENGTH <- as.factor(df_s$LENGTH)
df_s$GRAM <- as.factor(df_s$GRAM)
#class(df_s$LENGTH)
#class(df_s$GRAM)
#levels(df_s$LENGTH)
#levels(df_s$GRAM)
contrasts(df_s$LENGTH) <- c(0.5,-0.5)
contrasts(df_s$GRAM) <- c(-0.5,0.5)
```

```{r}
#V1
df_v_first$LENGTH <- as.factor(df_v_first$LENGTH)
df_v_first$GRAM <- as.factor(df_v_first$GRAM)
contrasts(df_v_first$LENGTH) <- c(0.5,-0.5)
contrasts(df_v_first$GRAM) <- c(-0.5,0.5)
#V2
df_v$LENGTH <- as.factor(df_v$LENGTH)
df_v$GRAM <- as.factor(df_v$GRAM)
contrasts(df_v$LENGTH) <- c(0.5,-0.5)
contrasts(df_v$GRAM) <- c(-0.5,0.5)
#O1
df_o_first$LENGTH <- as.factor(df_o_first$LENGTH)
df_o_first$GRAM <- as.factor(df_o_first$GRAM)
contrasts(df_o_first$LENGTH) <- c(0.5,-0.5)
contrasts(df_o_first$GRAM) <- c(-0.5,0.5)
#O2
df_o$LENGTH <- as.factor(df_o$LENGTH)
df_o$GRAM <- as.factor(df_o$GRAM)
contrasts(df_o$LENGTH) <- c(0.5,-0.5)
contrasts(df_o$GRAM) <- c(-0.5,0.5)
#A1.1
df_a1_first$LENGTH <- as.factor(df_a1_first$LENGTH)
df_a1_first$GRAM <- as.factor(df_a1_first$GRAM)
contrasts(df_a1_first$LENGTH) <- c(0.5,-0.5)
contrasts(df_a1_first$GRAM) <- c(-0.5,0.5)
#A1.2
df_a1$LENGTH <- as.factor(df_a1$LENGTH)
df_a1$GRAM <- as.factor(df_a1$GRAM)
contrasts(df_a1$LENGTH) <- c(0.5,-0.5)
contrasts(df_a1$GRAM) <- c(-0.5,0.5)
#A2.1
df_a2_first$LENGTH <- as.factor(df_a2_first$LENGTH)
df_a2_first$GRAM <- as.factor(df_a2_first$GRAM)
contrasts(df_a2_first$LENGTH) <- c(0.5,-0.5)
contrasts(df_a2_first$GRAM) <- c(-0.5,0.5)
#A2.2
df_a2$LENGTH <- as.factor(df_a2$LENGTH)
df_a2$GRAM <- as.factor(df_a2$GRAM)
contrasts(df_a2$LENGTH) <- c(0.5,-0.5)
contrasts(df_a2$GRAM) <- c(-0.5,0.5)
```
## Check of class and level
```{r}
class(df_v_first$LENGTH)
class(df_v_first$GRAM)
levels(df_v_first$LENGTH)
levels(df_v_first$GRAM)
class(df_v$LENGTH)
class(df_v$GRAM)
levels(df_v$LENGTH)
levels(df_v$GRAM)
class(df_o_first$LENGTH)
class(df_o_first$GRAM)
levels(df_o_first$LENGTH)
levels(df_o_first$GRAM)
class(df_o$LENGTH)
class(df_o$GRAM)
levels(df_o$LENGTH)
levels(df_o$GRAM)
class(df_a1_first$LENGTH)
class(df_a1_first$GRAM)
levels(df_a1_first$LENGTH)
levels(df_a1_first$GRAM)
class(df_a1$LENGTH)
class(df_a1$GRAM)
levels(df_a1$LENGTH)
levels(df_a1$GRAM)
class(df_a2_first$LENGTH)
class(df_a2_first$GRAM)
levels(df_a2_first$LENGTH)
levels(df_a2_first$GRAM)
class(df_a2$LENGTH)
class(df_a2$GRAM)
levels(df_a2$LENGTH)
levels(df_a2$GRAM)
```
## MODELS
## SUBJEKT
## First fixation duration (S)
```{r}
m.s.firstfix <- lmer (IA_FIRST_FIXATION_DURATION ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=df_s_first, REML=F)
summary(m.s.firstfix)
```
## Gaze duration (S)
```{r}
m.s.gaze <- lmer (IA_FIRST_RUN_DWELL_TIME ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=df_s_first, REML=F)
summary(m.s.gaze)
```
## Regression path duration (S)
```{r}
m.s.reg.path <- lmer (IA_REGRESSION_PATH_DURATION ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=df_s_first, REML=F)
summary(m.s.reg.path)
```
## Total duration (S)
Remember to change subset (first-pass reading)!
```{r}
m.s.total <- lmer (IA_DWELL_TIME ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=df_s, REML=F)
summary(m.s.total)
```
## Plots for article
```{r}
plot7 <- plot(effect('GRAM:TRIAL_INDEX_targets',m.s.reg.path))
plot7
```

```{r}
plot8 <- plot(effect('GRAM:TRIAL_INDEX_targets',m.s.total))
plot8
```
## Regressions out (S) - rescaled trial_index
```{r}
s.reg.out <- glmer (IA_REGRESSION_OUT ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_RESCALED + (1|identifier) + (1|PARTICIPANT), data=df_s_first, family = binomial)
summary(s.reg.out)
```
## Skipping percentage (S) - rescaled trial_index
```{r}
s.skip <- glmer (IA_SKIP ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_RESCALED + (1|identifier) + (1|PARTICIPANT), data=df_s, family = binomial)
summary(s.skip)
```
## VERB
## First fixation duration (V)
```{r}
m.v.firstfix <- lmer (IA_FIRST_FIXATION_DURATION ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=df_v_first, REML=F)
summary(m.v.firstfix)
```
## Gaze duration (v)
```{r}
m.v.gaze <- lmer (IA_FIRST_RUN_DWELL_TIME ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=df_v_first, REML=F)
summary(m.v.gaze)
```
## Regression path duration (v)
```{r}
m.v.reg.path <- lmer (IA_REGRESSION_PATH_DURATION ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=df_v_first, REML=F)
summary(m.v.reg.path)
```
## Total duration (v)
```{r}
m.v.total <- lmer (IA_DWELL_TIME ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=df_v, REML=F)
summary(m.v.total)
```
## Plots for article
```{r}
plot9 <- plot(effect('GRAM:TRIAL_INDEX_targets',m.v.reg.path))
plot9
```

```{r}
plot10 <- plot(effect('GRAM:TRIAL_INDEX_targets',m.v.total))
plot10
```
## Regressions out (v) - rescaled trial_index
```{r}
v.reg.out <- glmer (IA_REGRESSION_OUT ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_RESCALED + (1|identifier) + (1|PARTICIPANT), data=df_v_first, family = binomial)
summary(v.reg.out)
```
## Skipping percentage (v) - rescaled trial_index
```{r}
v.skip <- glmer (IA_SKIP ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_RESCALED + (1|identifier) + (1|PARTICIPANT), data=df_v, family = binomial)
summary(v.skip)
```
## ADVERB 1
## First fixation duration (A1)
```{r}
m.a1.firstfix <- lmer (IA_FIRST_FIXATION_DURATION ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=df_a1_first, REML=F)
summary(m.a1.firstfix)
```
## Gaze duration (a1)
```{r}
m.a1.gaze <- lmer (IA_FIRST_RUN_DWELL_TIME ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=df_a1_first, REML=F)
summary(m.a1.gaze)
```
## Regression path duration (a1)
```{r}
m.a1.reg.path <- lmer (IA_REGRESSION_PATH_DURATION ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=df_a1_first, REML=F)
summary(m.a1.reg.path)
```
## Total duration (a1)
```{r}
m.a1.total <- lmer (IA_DWELL_TIME ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=df_a1, REML=F)
summary(m.a1.total)
```
## Plot for article
```{r}
plot6 <- plot(effect('GRAM:TRIAL_INDEX_targets',m.a1.total))
plot6
```
## Skipping percentage (a1) - rescaled trial_index
Only converges without GRAM*TRIAL + optimizer
```{r}
a1.skip <- glmer (IA_SKIP ~ GRAM * LENGTH + TRIAL_INDEX_RESCALED + (1|identifier) + (1|PARTICIPANT), data=df_a1, family = binomial, control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))
summary(a1.skip)
```
## OBJEKT
## First fixation duration (o)
```{r}
m.o.firstfix <- lmer (IA_FIRST_FIXATION_DURATION ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=df_o_first, REML=F)
summary(m.o.firstfix)
```
## Gaze duration (o)
```{r}
m.o.gaze <- lmer (IA_FIRST_RUN_DWELL_TIME ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=df_o_first, REML=F)
summary(m.o.gaze)
```
## Regression path duration (o)
```{r}
m.o.reg.path <- lmer (IA_REGRESSION_PATH_DURATION ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=df_o_first, REML=F)
summary(m.o.reg.path)
```
## Plot for regression path duration (o)
```{r}
plot4 <- plot(effect('GRAM:LENGTH', m.o.reg.path))
plot4
```
## Total duration (o)
```{r}
m.o.total <- lmer (IA_DWELL_TIME ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=df_o, REML=F)
summary(m.o.total)
```
## Regressions out (o) - rescaled trial_index
```{r}
o.reg.out <- glmer (IA_REGRESSION_OUT ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_RESCALED + (1|identifier) + (1|PARTICIPANT), data=df_o_first, family = binomial)
summary(o.reg.out)
```
## Skipping percentage (o) - rescaled trial_index
```{r}
o.skip <- glmer (IA_SKIP ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_RESCALED + (1|identifier) + (1|PARTICIPANT), data=df_o, family = binomial)
summary(o.skip)
```
## ADVERB 2
## First fixation duration (A2)
```{r}
m.a2.firstfix <- lmer (IA_FIRST_FIXATION_DURATION ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=df_a2_first, REML=F)
summary(m.a2.firstfix)
```
## Gaze duration (a2)
```{r}
m.a2.gaze <- lmer (IA_FIRST_RUN_DWELL_TIME ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=df_a2_first, REML=F)
summary(m.a2.gaze)
```
## Regression path duration (a2)
```{r}
m.a2.reg.path <- lmer (IA_REGRESSION_PATH_DURATION ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=df_a2_first, REML=F)
summary(m.a2.reg.path)
```
## Total duration (a2)
```{r}
m.a2.total <- lmer (IA_DWELL_TIME ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=df_a2, REML=F)
summary(m.a2.total)
```
## Regressions out (a2) - rescaled trial_index
```{r}
a2.reg.out <- glmer (IA_REGRESSION_OUT ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_RESCALED + (1|identifier) + (1|PARTICIPANT), data=df_a2_first, family = binomial)
summary(a2.reg.out)
```
## Skipping percentage (a2) - rescaled trial_index
Does not converge with interactions
```{r}
#b.df_a2$LENGTH = relevel(as.factor(b.df_a2$LENGTH), ref = "short")
a2.skip <- glmer (IA_SKIP ~ GRAM + LENGTH + TRIAL_INDEX_RESCALED + (1|identifier) + (1|PARTICIPANT), data=df_a2, family = binomial)
summary(a2.skip)
```
## MODEL TOTAL READING TIME FOR THE WHOLE SENTENCE
```{r}
m.total <- lmer (DWELL_TIME_SENTENCE ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=df_s, REML=F)
summary(m.total)
```
## Individual variation (plot for article)
```{r}
plot3 <- augment(m.total) %>%
    ggplot(aes(GRAM,color = LENGTH, group = PARTICIPANT)) + facet_wrap(~LENGTH)+
    geom_jitter(aes(y = DWELL_TIME_SENTENCE), alpha = 0.5, height = 0, width = .1) +
    stat_summary(fun = mean, aes(y = .fitted), geom = "line", alpha = .5) + theme_classic()
plot3
```
## Looking at the individuals (plot for article)
```{r}
plot11 <- augment(m.total) %>%
    ggplot(aes(GRAM,color = LENGTH, group = LENGTH)) + facet_wrap(~PARTICIPANT)+
    geom_jitter(aes(y = DWELL_TIME_SENTENCE), alpha = 0.5, height = 1, width = .1) +
    geom_line(aes(y = .fitted)) + theme_classic()
plot11
```
## Plots for article
```{r}
plot1 <- plot(effect('GRAM:LENGTH', m.total))
plot1
```

```{r}
plot2 <- plot(effect('GRAM:TRIAL_INDEX_targets', m.total))
plot2
```
## ANALYSIS OF COLLAPSED REGIONS (S+V)
Collapsing the two regions
```{r}
subj_verb = subset(Eye, IA_ID == 2 | IA_ID == 3)
subj_verb = subj_verb %>%
  drop_na()  %>% 
  group_by(PARTICIPANT, TRIAL_INDEX_targets, TRIAL_INDEX_RESCALED, identifier, condition, GRAM, LENGTH, sentence) %>%

  summarize(IA_FIRST_FIX_PROGRESSIVE = max(as.numeric(IA_FIRST_FIX_PROGRESSIVE)),
            IA_FIRST_FIXATION_DURATION = sum(as.numeric(IA_FIRST_FIXATION_DURATION)),
            IA_FIRST_RUN_DWELL_TIME = sum(as.numeric(IA_FIRST_RUN_DWELL_TIME)),
            IA_REGRESSION_PATH_DURATION = sum(as.numeric(IA_REGRESSION_PATH_DURATION)),
            IA_DWELL_TIME = sum(as.numeric(IA_DWELL_TIME)),
            IA_REGRESSION_OUT = max(as.numeric(IA_REGRESSION_OUT)),
            IA_SKIP = max(as.numeric(IA_SKIP)),
            DWELL_TIME_SENTENCE = as.numeric(DWELL_TIME_SENTENCE)) 
subj_verb = subj_verb[!duplicated(subj_verb),]
write.csv(subj_verb, "subj_verb.csv", row.names = FALSE)
```
## Subsets and contrast coding for collapsed region
```{r}
# Subset during first-pass reading
subj_verb_first <- subset(subj_verb, IA_FIRST_FIX_PROGRESSIVE == "1")
```

```{r}
subj_verb$LENGTH <- as.factor(subj_verb$LENGTH)
subj_verb$GRAM <- as.factor(subj_verb$GRAM)
class(subj_verb$LENGTH)
class(subj_verb$GRAM)
levels(subj_verb$LENGTH)
levels(subj_verb$GRAM)
contrasts(subj_verb$LENGTH) <- c(0.5,-0.5)
contrasts(subj_verb$GRAM) <- c(-0.5,0.5)

subj_verb_first$LENGTH <- as.factor(subj_verb_first$LENGTH)
subj_verb_first$GRAM <- as.factor(subj_verb_first$GRAM)
class(subj_verb_first$LENGTH)
class(subj_verb_first$GRAM)
levels(subj_verb_first$LENGTH)
levels(subj_verb_first$GRAM)
contrasts(subj_verb_first$LENGTH) <- c(0.5,-0.5)
contrasts(subj_verb_first$GRAM) <- c(-0.5,0.5)
```
## First fixation duration (S+V)
```{r}
kol.firstfix <- lmer (IA_FIRST_FIXATION_DURATION ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=subj_verb_first, REML=F)
summary(kol.firstfix)
```
## Gaze duration (S+V)
```{r}
kol.gaze <- lmer (IA_FIRST_RUN_DWELL_TIME ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=subj_verb_first, REML=F)
summary(kol.gaze)
```
## Regression path duration (S+V)
```{r}
kol.reg.path <- lmer (IA_REGRESSION_PATH_DURATION ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=subj_verb_first, REML=F)
summary(kol.reg.path)
```
## Total duration (S+V)
```{r}
kol.total <- lmer (IA_DWELL_TIME ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_targets + (1|identifier) + (1|PARTICIPANT), data=subj_verb, REML=F)
summary(kol.total)
```
## Regressions out (S+V) - rescaled trial_index
```{r}
kol.reg.out <- glmer (IA_REGRESSION_OUT ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_RESCALED + (1|identifier) + (1|PARTICIPANT), data=subj_verb_first, family = binomial)
summary(kol.reg.out)
```
## Skipping percentage (S+V) - rescaled trial_index
```{r}
kol.skip <- glmer (IA_SKIP ~ GRAM * LENGTH + GRAM * TRIAL_INDEX_RESCALED + (1|identifier) + (1|PARTICIPANT), data=subj_verb, family = binomial)
summary(kol.skip)
```
## Plot for article - first fixation on the verb
```{r}
plot5 <- ggplot(df_v_first,aes(y=IA_FIRST_FIXATION_DURATION,x=TRIAL_INDEX_targets, color=GRAM)) +
  geom_point(aes(shape=GRAM))+
  geom_smooth(method='lm')+
  scale_color_manual(values = c("#4E84C4", "#D16103"))+
  facet_grid(~LENGTH)+
  ggtitle("First fixation duration on the verb")+
  ylab("First fixation duration (ms)")+
  xlab("Presentation order (from 1-40)")+
  labs(colour = "Condition")
plot5
```
